---
title: "Model Diagnostics"
output: 
  html_document:
    toc: true
    toc_float: true
    depth: 2
params:
  model_run: 106
  model_dir: "models/pk"
  script: "pk-diagnostics.Rmd"
  yspec: "data/derived/pk.yml"
  #yspec_namespace: ["plot", "tex"] # optionally uncomment to use alternate spec namespace
  cont_cov_flag: "diagContCov"
  cat_cov_flag: "diagCatCov"
  eta_names: ["ETA1//ETA-KA", "ETA2//ETA-V/F", "ETA3//ETA-CL/F"]
  log_dv: FALSE
  run_mrggsave: FALSE
---


```{r setup_general, include = FALSE, message=FALSE, warning = FALSE}
### General script set-up
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      height = 7, width = 6)

### Packages ----------------------------
library(here) 
library(tidyverse)
library(bbr)
library(pmtables)
library(pmplots)
library(mrggsave)
library(mrgmisc)
library(yaml)
library(yspec)
library(glue)
library(patchwork)

### Directories ----------------------------
figDir <- here("results", "figure", "diagnostics") 
if(!dir.exists(figDir)) dir.create(figDir)

### Source function scripts ----------------------------
source(here("scripts", "functions", "functions-diagnostics.R"))

### Model directory ----------------------------
thisModelPath <- file.path(params$model_dir, params$model_run)

### Figure location ----------------------------
thisFigDir <- file.path(figDir, params$model_run)
if(!dir.exists(thisFigDir)) dir.create(thisFigDir)

### Options ----------------------------
options(mrg.script = params$script, 
        mrggsave.dir = thisFigDir,   # set figure directory here
        mrggsave.dev = "pdf, png"         # save pdf, png, or both ("pdf, png")
        )
theme_set(theme_bw())
margin <- theme(plot.margin = unit(c(0.2, 0.2, 0.1, 0.1), "in"))
run_mrggsave <- params$run_mrggsave 


```

```{r read_mod_summary, results = "asis"}
### Read in model and generate summary ----------------------------
mod <- read_model(here(thisModelPath))
summ <- mod %>% model_summary()
```


# Summaries for model run: `r params$model_run`

### Quick look

- Dataset: `r fs::path_rel(summ$run_details$data_set, here())`
- Number of subjects: `r summ$run_details$number_of_subjects`
- Number of observations: `r summ$run_details$number_of_obs`
- Number of records: `r summ$run_details$number_of_data_records`
- Objective function value (final est. method): `r as.character(round(summ$ofv[[length(summ$ofv)]]$ofv_no_constant, 1))`
- Condition number: `r sig(summ$condition_number[[1]]$condition_number, maxeval=5)`
- Figures saved to: `r fs::path_rel(thisFigDir, here())`

Model description (from yaml file) is included below:

_`r mod$description`_


### Output summaries from bbr

```{r print_model_summary, results = "asis"}
### Print bbr model summaries in html ----------------------------
mod
summ
```

```{r yspec_details}
### Load Spec ----------------------------
# Information in the spec file is used to label the figures automatically.

# Check a spec was provided 
stopifnot(is.character(params$yspec))

# load yspec
spec <- ys_load(here(params$yspec)) 
  
# if namespaces provided in spec
if(is.character(params$yspec_namespace)){
  spec <- ys_namespace(spec, params$yspec_namespace)
}

### Define covariates  ----------------------------
# Requires the reader to provide flags in the spec
contCov <- pull_meta(spec, "flags")[[params$cont_cov_flag]]
catCov <- pull_meta(spec, "flags")[[params$cat_cov_flag]]

```


```{r read_in_data, message=FALSE}

# Read in model ----------------------------
mod <- read_model(here(thisModelPath))

# Join original dataset and nonmem output ----------------------------
data0 <- nm_join(mod)

# The NONMEM input dataset and output tables are read in and joined by a `NUM` 
# column (with bbr::nm_join). **This assumes that a row number column (called `NUM`) 
# was included during data assembly and is output in all $TABLE.** 
# The idea is that in NONMEM, you table just `NUM` and none of the other input data items. 
# They all get joined back to the NONMEM output (including character columns).

# If DV was estimated in the log domain ----------------------------
if(params$log_dv) {
  data0 <- data0 %>%
    mutate(DV = exp(DV),
           IPRED = exp(IPRED), 
           PRED = exp(PRED))  
}
# Note, this code assumes the user tabled DV, PRED, and IPRED in the NONMEM model

# Observation rows only and factors ----------------------------
data <- data0 %>%
  filter(EVID == 0) %>%
  ys_factors(spec)

# Note, the `data` used in the diagnostic plots has been filtered to only 
# include the observations (i.e., `EVID == 0`). Further modifications maybe needed, 
# for example, if BLQ data was included in the model or if the `DV` was log-transformed. 

# Single row per ID ----------------------------
id <- distinct(data, ID, .keep_all = TRUE)

# The `id` subset gets the first record per ID. This would usually be the 
# baseline value but consider filtering on a baseline flag if available. Also, 
# if the model includes inter-occasion variability (IOV), the occasion variable 
# should be included within the `distinct` function. 

```


# General diagnostic plots

```{r xaxis_label_names}
# Get label and units from spec (short and units)  ----------------------------
# Code currently assumes the time and tad columns named TIME and TAD
xTIME <- pm_axis_time(spec$TIME$unit)
xTAD <- pm_axis_tad(spec$TAD$unit)
xPRED <- pm_axis_pred(tolower(ys_get_short_unit(spec)$DV))
xDV <- ys_get_short_unit(spec)$DV
```

## DV vs PRED and IPRED

DV vs PRED and IPRED linear scale

```{r dv_pred_linear, message = FALSE}

dvp <- dv_pred(data, yname = xDV)     # DV vs PRED
dvip <- dv_ipred(data, yname = xDV)   # DV vs IPRED

p <- (dvp / dvip) + margin + plot_layout(axes = "collect")

```

```{r print_dv_pred_linear, echo = F, message = FALSE, fig.height = 7, fig.width = 5}
p
```

```{r save_dv_pred_linear, eval = run_mrggsave, include = FALSE, message = FALSE}
mrggsave_last(stem = "{params$model_run}-dv-pred-ipred-linear", height = 7)
```

```{r rm_dv_pred_linear, include = FALSE}
rm(p, dvp, dvip)
```

DV vs PRED and IPRED log scale

```{r dv_pred_log, message = FALSE}

dvp <- dv_pred(data, yname = xDV, loglog = TRUE) 
dvip <- dv_ipred(data, yname = xDV, loglog = TRUE)

p <- (dvp / dvip) + margin + plot_layout(axes = "collect")

```

```{r print_dv_pred_log, message = FALSE, fig.height = 7, fig.width = 5}
p
```

```{r save_dv_pred_log, eval = run_mrggsave, include = FALSE, message = FALSE}
mrggsave_last(stem = "{params$model_run}-dv-pred-ipred-log", height = 7)
```

```{r rm_dv_pred_log, include = FALSE}
rm(p, dvp, dvip)
```



## CWRES plots

CWRES vs PRED, time and time after dose.

```{r combined_cwres, message = FALSE}
p <- cwres_panel(data)
```

```{r print_combined_cwres, message = FALSE, fig.height = 7, fig.width = 5}
p
```

```{r save_combined_cwres, eval = run_mrggsave, include = FALSE, message = FALSE}
mrggsave_last(stem = "{params$model_run}-cwres-pred-time", height = 7)
```

```{r rm_combined_cwres, include = FALSE}
rm(p)
```


# EBEs-based diagnostics

## ETA pairs plot

```{r eta_pairs, message = FALSE}
p <- eta_pairs(id, params$eta_names)
```

```{r print_eta_pairs, message = FALSE, fig.height = 6, fig.width = 6}
p
```

```{r save_eta_pairs, eval = run_mrggsave, include = FALSE, message = FALSE}
mrggsave_last(stem = "{params$model_run}-eta-pairs", height = 6, width = 6)
```

```{r rm_eta_pairs, include = FALSE}
rm(p)
```

<!-- Covariate plots -->

```{r get_cont_cov_labels}
# get continuous cov details from spec  ----------------------------
co <- spec %>%
  ys_select(any_of(contCov)) %>%             
  axis_col_labs(title_case = TRUE,   
                short_max = 15)

# get categorical cov details from spec  ----------------------------
ca <- spec %>%
  ys_select(any_of(catCov)) %>%             
  axis_col_labs(title_case = TRUE,   
                short_max = 20) 

```

## ETA vs continous covariates


```{r eta_cont_cov, message = FALSE}

# Create plots with a single ETA and multiple covariates
p <- eta_covariate(id, x = co, y = params$eta_names)

# Create plots with a single covariate and multiple ETAs
# p <- eta_covariate(id, x = co, y = params$eta_names, transpose = TRUE)

p <- map(p, ~ .x + plot_layout(axes = "collect"))

```

```{r print_eta_cont_cov, message = FALSE, fig.height = 7, fig.width = 6}
 walk(p, print)
```

```{r save_eta_cont_cov, eval = run_mrggsave, include = FALSE, message = FALSE}
mrggsave(p, stem = "{params$model_run}-eta-cont-cov", height = 7, width = 6)
```

```{r rm_eta_cont_cov, include = FALSE}
rm(p)
```

## ETA vs categorical covariates

```{r eta_cat_cov, message = FALSE}

# Create plots with a single ETA and multiple covariates
p <- eta_covariate(id, x = ca, y = params$eta_names, ncol = 1)

# Create plots with a single covariate and multiple ETAs
# p <- eta_covariate(id, x = ca, y = params$eta_names, transpose = TRUE)

p <- map(p, ~ .x + plot_layout(axes = "collect"))

```

```{r print_eta_cat_cov, message = FALSE, fig.height = 6, fig.width = 6}
walk(p, print)
```

```{r save_eta_cat_cov, eval = run_mrggsave, include = FALSE, message = FALSE}
mrggsave(p, stem = "{params$model_run}-eta-cat-cov", height = 6, width = 6)
```

```{r rm_eta_cat_cov, include = FALSE}
rm(p)
```


# Session details

```{r details, include = TRUE}
date()
sessionInfo()
bbi_version()
```

