---
title: "Model Management - Working Notebook"
editor_options: 
  chunk_output_type: console
---

**NOTE: This `model-management.Rmd` file is meant to be a fluid "working
notebook"** for the Pharmacometrician to use throughout the project. Some points
to consider:

* This file is _not_ meant to "knit" to HTML or PDF.

* This file will _not_ be submitted for QC at any point. QC of the relevant
activities will be done via other scripts (e.g. `model-overview.R`, `run-log.R`,
etc.).

* You do _not_ need to copy/paste code in here in order to preserve the story of
your modeling activities. Instead, re-use code chunks as needed, changing the
name of model as you iterate, and updating any specific pieces of code where
relevant.

* Attach notes, description, and other annotation to your model object as you go
(via `bbr`'s `add_notes()`, `add_description()`, etc.). **`bbr` will keep track
of this information for you. You do _not_ need to keep these calls (or the
associated notes) in this script to preserve them.**

* All notes, description, etc. can be accessed later by re-loading the specific
model, or by viewing them in the tibble returned from `run_log()`.


```{r setup}
library(bbr)
library(tidyverse)
library(here)
library(pmplots)
library(magrittr)
library(yspec)

source(here("scripts/functions/bbr-helpers.R"))
source(here("scripts/functions/functions-diagnostics.R"))
source(here("scripts/functions/model-check-helpers.R"))

# define model dir and load yspec YAML
modDir <- here("models/pk")
spec_path <- here("data/derived/pk.yml")
spec <- load_spec(spec_path)

# Install bbi (once per user per project)
# bbr::use_bbi(.path = getOption("bbr.bbi_exe_path"))

# Initialize bbi (once per modeling directory)
# bbi_init(modDir, .nonmem_dir = "/opt/NONMEM", .nonmem_version = "nm75")
```


# 1 run log

```{r, eval=FALSE}
rl_cols <- c("run","description","based_on","ofv", "dofv", "aic","notes", "error_msg")
rl <- run_log(modDir) %>% 
  add_config() %>% suppressWarnings() %>%
  add_summary() %>%
  add_dofv() %>%
  collapse_to_string(notes) %>%
  select(c(all_of(rl_cols)), contains("has_changed"))

View(rl, "runlog")
```

# 2 create models

## first model

```{r, eval=FALSE}
# Use this when starting a new model from the beginning
#  i.e. a model that is _not_ based on an existing model

# Path to control stream file
#  * Replace "<model-name>.ctl" with name of your control stream file
ctl_path <- here::here("models", "pk", "<model-name>.ctl") 

# Optionally, uncomment this to copy an example control stream file, as a starting point
# fs::file_copy(
#   here::here("models", "pk", "102.ctl"),
#   ctl_path
# )

if (!fs::file_exists(ctl_path)) warning(paste("Please copy control stream file to", ctl_path))

# Create a bbr model object from this control stream file
mod <- new_model(
  fs::path_ext_remove(ctl_path),
  # .tags = "<my tag>" # uncomment to add tags
)

# open file for editing in RStudio
open_model_file(mod) 
```


## model iteration

```{r, eval=FALSE}
PARENT_MODEL <- 106 # Name of most recently completed model
THIS_MODEL <- 107   # Name of new model to create (based on PARENT_MODEL)

set.seed(240321) # for reproducibility of tweak_param_estimates(), below

parent_mod <- read_model(file.path(modDir, PARENT_MODEL))

mod <- copy_model_from(
  .parent_mod=parent_mod, 
  .new_model=THIS_MODEL, # will increment, or pass custom name if desired
  # .overwrite = TRUE,   # uncomment this line to overwrite existing THIS_MODEL
  # .inherit_tags = TRUE # uncomment this line to inherit any tags from parent model
) %>%
  update_model_id() %>%
  inherit_param_estimates() %>%
  tweak_initial_estimates()

# open file for editing in RStudio
open_model_file(mod) 
```

## add model annotation
Annotating your models as you work enhances traceability and reproduceability.
You do _not_ have to maintain the code for these model annotation calls; `bbr`
keeps track of all annotations for later reference. They can then be used in Run
Logs and other reporting outputs, as needed.

Also note that **these annotation functions can be run at any time, before or
after executing your model.**

Consider the following guidance:

* **`description`** is used to document the basic model structure. 
* What you add here will be included in the final Run Log table. It should 
"stand alone" to describe model structure. 
* You will need to enter this _manually, for each new model_.

* **`notes`** are for informal notes on decision points, model performance, etc. 
* These are for later reference by the modeler or collaborators, and will 
_not_ be included in the final Run Log. 
* Often, `notes` are used to describe notable changes from parent model. 

* **`star`** only your key models. 
* Only starred models will be included in the final Run Log table in the report. 
* You can use `add_star(mod)` or `remove_star(mod)` at any point in the 
analysis process.

* **`tags`** are optional organizational helpers. 
* These are often used to organize "categories" of models. 
* You can easily filter a Run Log to only models with a given tag by using 
`run_log(..., .include = "<my tag>")`.


```{r}
# preview the description of the parent model
print(parent_mod$description)
```

```{r}
# uncomment the next line and fill in description of new model 
#   to describe the basic model structure
# mod <- add_description(mod, "<description of new model>")
print(mod$description)
```


```{r}
# optionally, add notes or tags before executing
# mod <- mod %>% add_notes("<my note>")
# mod <- mod %>% add_tags("<my tag>")
```

```{r}
## compare control stream to parent model
model_diff(mod)
```


# 3 execute model

```{r, eval=FALSE}
# set default parallelization for this model
#   see bbr::print_bbi_args() for other options that can be set
mod <- mod %>% add_bbi_args(
  list(parallel = TRUE, threads = 8) 
)
```

## submit local

```{r, eval=FALSE}
check_model_setup(mod, .join_col = "NUM")
```
```{r}
submit_model(mod) # submit model to run on compute nodes
# submit_model(mod, .mode="local", .wait = FALSE) # or submit model to run on head node  
```

## monitoring

```{r, eval = FALSE}
# check if run has finished
check_nonmem_finished(mod)
# wait_for_nonmem(mod) # optionally hold the R console until model finishes

# print head and tail of output files to R console
tail_lst(mod)
tail_output(mod)

# open .lst file to view in RStudio
build_path_from_model(mod, ".lst") %>% file.edit()

# open OUTPUT or PRDERR files to view in RStudio
open_file_if_exists(mod, "OUTPUT")
open_file_if_exists(mod, "PRDERR")
```

## Simulate
```{r, eval=FALSE}
SIMULATE <- FALSE

if (isTRUE(SIMULATE)) {
  # Model to simulate
  smod <- read_model(file.path(modDir, THIS_MODEL))
  check_msf_file(smod)
  
  # Data
  data <- nm_data(smod)
  data <- filter(data, is.na(C))
  data <- mutate(data, BLQ = 0)
  
  # Simulate
  add_simulation(
    smod, 
    n = 300,  
    data = data, 
    sim_cols = c("DV", "PRED"), 
    .mode = "local",
    .join_col = "NUM", .wait = TRUE, .overwrite = TRUE,
    .bbi_args = list(threads = 1), # change this if you want more
    seed = 1234
  )
  
  sim <- get_simulation(smod)
  model_diff(sim) # Inspect differences of new control stream

  # Check output if finished
  if(check_nonmem_finished(smod)) {
    sims <- nm_join_sim(smod)
    nrow(sims)/nrow(data)
  }
}
```


# 4 view run results

## model summary

```{r, eval=FALSE}
modsum <- model_summary(mod) 
print(modsum)
```

```{r, eval=FALSE}
# preview final parameter estimates
param_df <- param_estimates(modsum) 
View(param_df)
```


## ad-hoc diagnostics

```{r, eval=FALSE}
can_be_nm_joined(mod)

# join input data to output tables
dat <- nm_join(mod) %>%
  dplyr::filter(EVID == 0) %>% # Filter to only observation records
  yspec::ys_add_factors(spec, .suffix = "") # Add factors to data set based on spec information
View(dat)

# make some simple diagnostic plots with MetrumRG's pmplots package
dv_pred(dat, yname = ys_get_short_unit(spec)$DV)
dv_ipred(dat, yname = ys_get_short_unit(spec)$DV)
cwres_panel(dat)
```

## render diagnostics templates

```{r, eval=FALSE}
# Render diagnostics with a parameterized Rmd template
#   cmd+click on function name to see documentation and argument options
model_diagnostics(
  .mod = mod,
  .p = list(
    cont_cov_flag = "diagContCov", 
    cat_cov_flag = "diagCatCov",
    eta_names = c("ETA1//ETA-KA", "ETA2//ETA-V/F", "ETA3//ETA-CL/F"),
    run_mrggsave = TRUE ),
  template =  here("scripts", "templates", "pk-diagnostics.Rmd")
) %>%
  browseURL()     # this will open the HTML doc in your browser

```

```{r, eval = FALSE}
model_diagnostics(
  .mod = mod,
  # .p = list(log_dv = TRUE),   ## optionally pass in document parameters
  template = here("scripts", "templates", "id-dv-pred-plots.Rmd")
) %>%
  browseURL()

```


## compare two models

```{r, eval=FALSE}
model_diff(mod) # compare to parent model
```


## add post-execution model annotation

These annotation functions can be run at any time, before or after executing 
or evaluating your model.

Review of relevant guidance from above:

* **`notes`** are for informal notes on decision points, model performance, etc. 
* These are for later reference by the modeler or collaborators, and will 
_not_ be included in the final Run Log. 
* Often, `notes` are used to describe notable changes from parent model. 

* **`star`** only your key models. 
* Only starred models will be included in the final Run Log table in the report. 
* You can use `add_star(mod)` or `remove_star(mod)` at any point in the 
analysis process.


```{r}
# uncomment to add any notes 
# * key decision points
# * notes on model performance
# * notable changes from parent model
# mod <- add_notes(mod, "<notes on new model>")

# IF this is a key model that we want to include in the report, 
#   uncomment the next line to add a "star" to it
# mod <- add_star(mod)
```


# 5 assorted utility code

## run many models
```{r, eval=FALSE}
mods_to_run <- 106:200
map(mods_to_run, ~read_model(file.path(modDir,.x))) %>% 
  submit_models(.bbi_args=list(overwrite=TRUE, parallel = TRUE, threads = 8))
```

## apply tag to many models
```{r, eval=FALSE}
mods_to_tag <- 106:200
walk(mods_to_tag, ~read_model(file.path(modDir,.x)) %>% 
       add_tags("<my tag>"))
```

## open many models in RStudio
```{r, eval=FALSE}
mods_to_open <- 106:200
walk(mods_to_open, ~read_model(file.path(modDir,.x)) %>% 
       open_model_file())
```
